Resonant Scattering in a Uniform Sphere with Large Optical Depth

### 2021-11-16 ms.tex
Phil's final changes up to eq 25

The outer layers of the atmosphere are central to a planet's evolution, since they can shelter the lower atmosphere from high energy radiation as well as regulate the escape of gas into space. Given the abundance of hydrogen and its large cross section, Lyman $\alpha$ (Ly$\alpha$) is an important component of the radiation field. There are two sources of Ly$\alpha$: the star, and recombinations in the planet's atmosphere. \lya may ionize atoms and dissociate molecules, as well as exert pressure forces that drive an outflow. \lya can also excite H atoms to the 2p state, creating a population of Balmer-line absorbers that can be observed via transmission spectroscopy \citep{2017ApJ...851..150H}. Due to the low gas densities in the upper atmosphere, collisional de-excitation and broadening are of secondary importance and \lya may undergo ``resonant scattering''.

Hubble Space Telescope (HST) observations with the STIS have found large \lya transit depths around a handful of exoplanets \citep{2003Natur.422..143V, 2012A&A...543L...4L, 2012A&A...547A..18E, 2015Natur.522..459E,  2017A&A...597A..26B, 2017A&A...599L...3B, 2017A&A...602A.106B, 2018A&A...620A.147B, 2019AJ....158...50W, 2019EPSC...13.1928L, 2020ApJ...888L..21G,2021arXiv210309864B}. These observations have revealed a population of atoms extending out to distances of order a few planetary radii for several planets around bright, nearby stars, motivating a study of the physics of \lya interactions with the H atom population. The transition from the atomic to the molecular layer in these atmospheres may take place at pressures of order ${\sim}\ 10\ \mu$bar (see discussion in \citet{2017ApJ...851..150H} for details). This suggests the presence of a thick layer of atomic H which can have a line center optical depth of $\tau_0\ {\sim}\ 10^8$. A careful treatment of resonant scattering is necessary in order to construct accurate models of H atom excitation, heating, and radiative forces.

Due to the technical challenge of including resonant scattering, the fully three-dimensional geometry, and the presence of an outflow, numerical simulations may be required to understand the dynamics of these irradiated exoplanet atmospheres. The extreme optical depths at \lya line center impose a large computational cost for solving radiative transfer with Monte Carlo methods directly coupled with fluid dynamics. The number of scatterings a photon undergoes is proportional to the line center optical depth, $\tau_0$, of the domain.  Near the base of the atomic layer, the line center optical depth is ${\sim}10^8$, so most of the time is spent following photons in these cells. A method that can treat these zones incorporating a direct solution rather than traditional Monte Carlo has the potential to greatly accelerate the calculation.

Approximate analytic solutions for resonant scattering exist in certain limits. \citet{1973MNRAS.162...43H} showed that when most of the radiation is out on the damping wings, the transfer equation reduces to the Poisson equation. However, their solution contains an incorrect treatment of the boundary condition. To our knowledge, the errors introduced by this treatment have never been quantified. They attempt a separation of variables $J(\tau,\sigma) = \theta(\tau) j(\sigma)$ in spatial variable $\tau$ and frequency variable $\sigma$ (their Equations 16 and 23). The solutions for the eigenfunctions $\theta(\tau)$ and $j(\sigma)$ then depend explicitly on the separation constant $\lambda$. In order to satisfy the boundary conditions, the separation constant is shown to satisfy an eigenvalue equation of the form
\be
\lambda \tan(\lambda B) & = & \frac{3}{2} \phi \Delta,
\label{eq:evalue}
\ee
where $2B$ is the slab optical depth at line center, $\phi$ is the line profile, and $\Delta$ is the Doppler width. The key point is that the line profile depends on one of the coordinates: frequency. This causes the eigenvalues of the separation constant to be frequency-dependent. Thus, the separation ``constant'' is not constant, and the function does not satisfy the Poisson equation since the frequency derivatives will act on the separation ``constant", giving extra terms. In the limit of large optical depth $B$, they approximate the eigenvalues as $\lambda_n B \simeq \pi (n-1/2)$. However, this gives zero mean intensity at the surface. Their Equation 34 subsequently allows $\lambda$ to have a small deviation from the above expression, which is explicitly frequency dependent. This allows a nonzero intensity at the surface, but at the cost of rendering the separation of variables assumption invalid. Our treatment, using the correct boundary condition, quantifies the errors in their ansatz.

@@

A solution for the mean intensity $J_d$ which is divergent at $\vec{x}=\vec{x_s}$ and $\sigma=\sigma_s$ and zero at infinity is presented in \citet{1990ApJ...350..216N}. Here it is extended to spherical geometry and generalized to allow emission frequencies away from line center: 
\be
J_{\rm d} & = & 
\left(\frac{\sqrt{6}k^2L}{16\pi^3 \Delta^3}\right)\left(\frac{1}{(kr/\Delta)^2 + (\sigma - \sigma_{\rm s})^2}\right)
\label{eq:Jd}
\ee

@@

We now estimate the scaling of $H_{\rm bc}$ with $\tau_0$. In the limit $J_{\rm bc} \gg H_{\rm bc}$, we find that $J_{\rm bc}\ {\sim}\ H_0$ from Equation \ref{eq:bc2}. We  calculate $H_{\rm bc}$ from $J_{\rm bc}$ using Equation \ref{eq:Hbc} as
\be
H_{\rm bc}(R, \sigma)\ {\sim}\ \frac{1}{3k\phi}\frac{ks}{\Delta}H_0\ {\sim}\ H_0 s \frac{x^2}{a}\ {\sim}\ H_0 \frac{1}{\tau_0}\frac{(a\tau_0)^{2/3}}{a}\ {\sim}\ H_0 (a\tau_0)^{-1/3},
\ee
where we have used $s\ {\sim}\ 1/\sigma\ {\sim}\ 1/\tau_0$ so that
\be \label{eq:hbc_scaling}
\frac{H_{\rm bc}(R, \sigma)}{H_0(R, \sigma)} \propto (a\tau_0)^{-1/3}.
\ee
At large $\tau_0$, it is expected that the correction term will be small, but it will become increasingly important as $\tau_0$ decreases. Our solution of the transfer equation is only valid when the peaks of the spectral energy distribution lie well outside of the Doppler core, e.g., for large $\tau_0$. For \lya and H atoms with $T=10^4$ K, the damping parameter is $a = 4.72\times 10^{-4}$. Thus, the value of $x$ at which the Doppler and Lorentzian components of the line wing are equal is $x_{\rm cw}=3.3$. Setting $x_{\rm cw} = x_{\rm peak}$ and solving for $\tau_0$ gives the value at which the peak of the spectrum falls at the Doppler core boundary, which is $\tau_{\rm cp} \approx 10^5$. Hence $H_{\rm bc}/H_0 {\sim} (\tau_{\rm cp}/\tau_0)^{1/3}$ is large at $\tau_0 \leq \tau_{\rm cp}$ and decreases relatively slowly as $\tau_0$ increases. Additionally, the optical depth at $x_{\rm peak}$ is proportional to $(a\tau_0)^{1/3}$, so photons here become optically thin when $a\tau_0 {\sim} 1$.

@@

\begin{figure}
    \centering
    \includegraphics{final_residual.pdf}
    \caption{Analytic solution for the outgoing spectrum compared with Monte Carlo at a line center optical depth of $\tau_0 = 1 \times 10^7$. Photons are initialized with $\rm x = 0.0$. In the log-scale plot in the second panel, $|\rm H_{bc}|$ is shown instead of $\rm H_{bc}$ to examine the contributions of both positive and negative pieces. Residuals of each solution to the Monte Carlo data points are shown in the third panel.} 
    \label{fig:sol_mc_residual_0}
\end{figure}

We now compare each of the previously-discussed solutions for surface flux to the Monte Carlo results. We return to the Doppler units $x$ for the photon frequency rather than the $\sigma$ units used in the solutions of the transfer equation. The spectrum $P(x)$ is defined as the specific luminosity at the surface divided by the source luminosity, or
\be
P(x) = \frac{16\pi^2R^2H(R, x)\Delta}{L}.
\ee
This is normalized such that $\int P(x)dx = 1$. Since $H(r, x)$ is per $d\nu$, a factor of $\Delta$ gives the expression the correct units.

@@

The size of $H_{\rm bc}$ is dependent on $\tau_0$, which also dictates the x-position of the peak of spectrum according to Eq. \ref{eq:tau_peak_scaling}. $H_{\rm bc}$ is significant even at $\tau_0 {\sim} 10^7$ where the $H_0$ solution is expected to perform well, i.e., photons are pushed further out into the wing where the simplifying assumptions made in the derivation of the differential equation are a better approximation.

In Figure \ref{fig:sol_mc_tau} we show the solutions alongside Monte Carlo for $\tau_0=10^5, 10^6$, and $10^7$. From Eq. \ref{eq:hbc_scaling}, the size of the term $H_{\rm bc}$ should become smaller with larger optical depths, following a $(a\tau_0)^{-1/3}$ scaling. Indeed, agreement between Eq. \ref{eq:H0surf} and the Monte Carlo points in Fig. \ref{fig:sol_mc_tau} improves as $\tau_0$ increases, with $H_{\rm bc}$ providing a fractionally smaller correction to $H_0$. One factor of $(a\tau_0)^{1/3}$ has been scaled out of the x-axis such that the peaks of the distributions are vertically aligned. This scaling has also been applied to the y-axis to preserve normalization of the escape probability. At lower $\tau_0$, the scattering of photons within the Doppler core of the line becomes important, but our analytic solution does not include this effect. The effects of line core scattering can be seen in the Monte Carlo data for $\tau_0=10^5$ and, to a lesser extent, $
\tau_0=10^6$.

 \begin{figure}
    \centering
    \includegraphics{xinit_threepanel.pdf}
    \caption{Analytic solution for the outgoing spectrum compared with Monte Carlo at a line center optical depth of $\tau_0 = 1 \times 10^7$. Photons are initialized with $\rm x = 0.0, 6.0, 12.0$ to examine the solutions' accuracy with a changing source frequency. The optical depth at each of these source frequencies is $\tau_s = 10^7, 77,$ and $19$, respectively.} 
    \label{fig:sol_mc_xinit}
\end{figure}

 \begin{figure}
    \centering
    \includegraphics{xinit_threepanel_tau1e6.pdf}
    \caption{Again changing $x_s$, but at a lower optical depth of $\tau_0 = 10^6$. The shift in $x$ is a much larger fraction of the distance to the spectral peak $(a\tau_0)^{1/3}$, and thus the bias in the spectrum is much more drastic. The optical depth at the source frequency is $\tau_s = 10^6$, $7.7$, and $1.9$ for $x_s=0.0, 6.0,$ and $12.0$, respectively. } 
    \label{fig:sol_mc_xinit_lowtau}
\end{figure}

@@

Figure \ref{fig:sol_mc_xinit} shows calculations performed for three initial photon frequencies $\rm x_s = 0.0, 6.0$, and $12.0$ for a line center optical depth of $\tau_0=10^7$. The asymmetry of the spectrum is slight for $\rm x_s = 6.0$ where the cross section is still relatively large, but the increase in mean free path is much more substantial when photons are initialized with $\rm x_s=12.0$, outside the line core. It is seen here that the difference between the Monte Carlo data and $H_0$ becomes larger as $\rm x_s$ increases. Thus, for sufficiently large $\tau_0$ and emission away from line center ($x_s \neq 0$), the importance of enforcing frequency-dependent boundary conditions at the surface of the sphere grows.

Figure \ref{fig:sol_mc_xinit_lowtau} shows emission away from line center at the same values of $x_s$ as in Figure \ref{fig:sol_mc_xinit}, but for $\tau_0=10^6$ rather than $10^7$. The difference between the left and right side of the escaping spectrum is substantial at lower optical depths, since $(a\tau_0)^{-1/3}$ increases by a factor of ${\sim}2$. It is clear from the figure that as $x_s$ extends further into the wing, the spectrum becomes more strongly peaked in frequency. Additionally, since the sphere is increasingly optically thin in the wing, we expect there to be a stronger disagreement between the Monte Carlo and the analytic solution as the equation assumed large optical depths.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{xinit.pdf}
    \caption{$H_{\rm bc}$ solutions shown at three line center optical depths, $\tau_0=10^5, 10^7,$ and $10^9$, for photon source frequencies $\sigma_s=0.0, \tau_0, 2\tau_0$. Thicker lines indicate larger optical depth, and darker lines indicate $\sigma_s$ further from line center. The axes are scaled to show that the size of the corrective factor agrees with the predicted $(a\tau_0)^{-1/3}$ scaling.}
    \label{fig:xinit}
\end{figure}

Figure \ref{fig:xinit} shows how the boundary term $H_{\rm bc}$ scales with both $\tau_0$ and $x_s$. In this figure, $x_s$ is shifted by integer multiples of $\tau_0$ in $\sigma$ frequency units (Eq. \ref{eq:change_of_variables}) such that the source, now written $\sigma_s$, falls near the peak of the spectrum for each $\tau_0$. For clarity, only the $x > 0$ side of the spectrum is shown. From Eq. \ref{eq:hbc_scaling}, it is expected that the fractional size of $H_{\rm bc}$ relative to $H_0$ should become smaller with larger optical depths following $(a\tau_0)^{-1/3}$. This factor has been scaled out of the figure such that solutions for different optical depths and the same source frequency $\sigma_s$ should show close agreement in scale on the figure's vertical axis if the relation holds. Indeed, the scaled solutions converge as the optical depth becomes larger, indicating general agreement with the $(a\tau_0)^{-1/3}$ scaling. The remaining discrepancy present in the vertical axis for fixed $\sigma_s$ results from $x_{\rm peak}$ becoming close to $x_{\rm cw}$; at lower $\tau_0$, this causes the line profile approximation in the wing (Eq. \ref{eq:app:line_profile_wing}) to break down. From this, we conclude that the errors introduced by the incorrect separation of variables in \cite{1973MNRAS.162...43H}, \cite{1990ApJ...350..216N}, \cite{2006ApJ...649...14D} and others are proportional to $(a\tau_0)^{-1/3}$.

@@

In order to understand how long it takes for the photons to escape the uniform sphere of gas we must reintroduce the time dependence of the diffusion equation, which was ignored in the steady-state calculations in Section \ref{sec:steadystate}. To obtain the radiative intensity $I=dE/(dAdtd\Omega d\nu)$ of a radiation field that is dynamic on timescales comparable to the light-crossing time $t_{\rm lc} = R/c$, the time-dependent response to a delta function impulse is found. This allows the distribution of photon escape times (the ``wait time distribution'') to be characterized. For simplicity a $J=0$ boundary condition will be used in the following derivations, which is a rough approximation for $a\tau_0 \gg 1$.

>>>>

The outer layers of the atmosphere are central to a planet's evolution, since they can shelter the lower atmosphere from high energy radiation as well as regulate the escape of gas into space. Given the abundance of hydrogen and its large cross section, the Lyman $\alpha$ (Ly$\alpha$) line is an important component of the radiation field. There are two sources of Ly$\alpha$: the star, and recombinations in the planet's atmosphere. \lya may ionize atoms and dissociate molecules, as well as exert pressure forces that drive an outflow. \lya can also excite H atoms to the 2p state, creating a population of Balmer-line absorbers that can be observed via transmission spectroscopy \citep{2017ApJ...851..150H}. Due to the low gas densities in the upper atmosphere, collisional de-excitation and broadening are of secondary importance and \lya may undergo ``resonant scattering''.

Hubble Space Telescope (HST) observations with the STIS have found large \lya transit depths around a handful of exoplanets \citep{2003Natur.422..143V, 2012A&A...543L...4L, 2012A&A...547A..18E, 2015Natur.522..459E,  2017A&A...597A..26B, 2017A&A...599L...3B, 2017A&A...602A.106B, 2018A&A...620A.147B, 2019AJ....158...50W, 2019EPSC...13.1928L, 2020ApJ...888L..21G,2021arXiv210309864B}. These observations have revealed a population of atoms extending out to distances of order a few planetary radii or more for several planets around bright, nearby stars, motivating a study of the physics of \lya interactions with the H atom population. The transition from the atomic to the molecular layer in these hot upper atmospheres may take place at pressures of order ${\sim}\ 10\ \mu$bar (see discussion in \citet{2017ApJ...851..150H} for details). This suggests the presence of a thick layer of atomic H which can have a line center optical depth of $\tau_0\ {\sim}\ 10^8$. A careful treatment of resonant scattering is necessary in order to construct accurate models of H atom excitation, heating, and radiative forces.

Due to the technical challenge of including resonant scattering, the fully three-dimensional geometry, and the presence of an outflow, numerical simulations may be required to fully understand the dynamics of these irradiated exoplanet atmospheres. The large optical depths at \lya line center impose a steep computational cost for solving radiative transfer with Monte Carlo methods directly coupled with fluid dynamics. The number of scatterings a photon undergoes is proportional to the line center optical depth, $\tau_0$, of the domain.  Near the base of the atomic layer, the line center optical depth is ${\sim}10^8$, so most of the time is spent following photons in these cells. A method that can treat these zones incorporating a direct solution rather than traditional Monte Carlo has the potential to greatly accelerate the calculation.

Approximate analytic solutions for resonant scattering exist in certain limits. \citet{1973MNRAS.162...43H} showed that when most of the radiation is out on the damping wings, the transfer equation reduces to the Poisson equation. However, their solution contains an incorrect treatment of the boundary condition. To our knowledge, the errors introduced by this treatment have never been quantified. They attempt a separation of variables as $J(\tau,\sigma) = \theta(\tau) j(\sigma)$ in spatial variable $\tau$ and frequency variable $\sigma$ (their Equations 16 and 23). The solutions for the eigenfunctions $\theta(\tau)$ and $j(\sigma)$ then depend explicitly on the separation constant $\lambda$. In order to satisfy the boundary conditions, the separation constant is shown to satisfy an eigenvalue equation of the form
\be
\lambda \tan(\lambda B) & = & \frac{3}{2} \phi \Delta,
\label{eq:evalue}
\ee
where $2B$ is the slab optical depth at line center, $\phi$ is the line profile, and $\Delta$ is the Doppler width. The key point is that the line profile depends on one of the coordinates: frequency. This causes the eigenvalues of the separation constant to be frequency-dependent. Thus, the separation ``constant'' is not constant, and the function does not satisfy the Poisson equation since the frequency derivatives will act on the separation ``constant", giving extra terms. In the limit of large optical depth $B$, they approximate the eigenvalues as $\lambda_n B \simeq \pi (n-1/2)$. However, this gives zero mean intensity at the surface. Their Equation 34 subsequently allows $\lambda$ to have a small deviation from the above expression, which is explicitly frequency dependent. This allows a nonzero intensity at the surface, but at the cost of rendering the separation of variables assumption invalid. Our treatment, using the correct boundary condition, quantifies the errors in their ansatz.

@@

A solution for the mean intensity $J_d$ which is divergent at $\vec{x}=\vec{x_s}=0$ %TODO: This ok?
and $\sigma=\sigma_s$ and zero at infinity is presented in \citet{1990ApJ...350..216N}. Here it is extended to spherical geometry and generalized to allow emission frequencies away from line center: 
\be
J_{\rm d} & = & 
\left(\frac{\sqrt{6}k^2L}{16\pi^3 \Delta^3}\right)\left(\frac{1}{(kr/\Delta)^2 + (\sigma - \sigma_{\rm s})^2}\right)
\label{eq:Jd}
\ee

@@

We now estimate the scaling of $H_{\rm bc}$ with $\tau_0$. In the limit $J_{\rm bc} \gg H_{\rm bc}$, we find that $J_{\rm bc} \approx \sqrt{3} H_0$ from Equation \ref{eq:bc2}. We  calculate $H_{\rm bc}$ from $J_{\rm bc}$ using Equation \ref{eq:Hbc} as
\be
H_{\rm bc}(R, \sigma) \approx \frac{1}{\sqrt{3}k\phi}\frac{ks}{\Delta}H_0\ {\sim}\ H_0 s \frac{x^2}{a}\ {\sim}\ H_0 \frac{1}{\tau_0}\frac{(a\tau_0)^{2/3}}{a}\ {\sim}\ H_0 (a\tau_0)^{-1/3},
\ee
where we have used $s\ {\sim}\ 1/\sigma\ {\sim}\ 1/\tau_0$ so that
\be \label{eq:hbc_scaling}
\frac{H_{\rm bc}(R, \sigma)}{H_0(R, \sigma)} \propto (a\tau_0)^{-1/3}.
\ee
At large $\tau_0$, it is expected that the correction term will be small, but it will become increasingly important as $\tau_0$ decreases. Our solution of the transfer equation is only valid when the peaks of the spectral energy distribution lie well outside of the Doppler core, e.g., for large $\tau_0$. For \lya and H atoms with $T=10^4$ K, the damping parameter is $a = 4.72\times 10^{-4}$. Thus, the value of $x$ at which the Doppler and Lorentzian components of the line wing are equal is $x_{\rm cw}=3.3$. Setting $x_{\rm cw} = x_{\rm peak}$ and solving for $\tau_0$ gives the value at which the peak of the spectrum falls at the Doppler core boundary, which is $\tau_{\rm cp} \approx 10^5$ (``core-peak'' optical depth). Hence $H_{\rm bc}/H_0 {\sim} (\tau_{\rm cp}/\tau_0)^{1/3}$ is large at $\tau_0 \leq \tau_{\rm cp}$ and decreases relatively slowly as $\tau_0$ increases. Additionally, the optical depth at $x_{\rm peak}$ is proportional to $(a\tau_0)^{1/3}$, so photons here become optically thin when $a\tau_0 {\sim} 1$.

@@

\begin{figure}
    \centering
    \includegraphics{final_residual.pdf}
    \caption{Spectrum $P(x)$ vs. frequency $x$. The analytic solutions for the outgoing spectrum are calculated at a line center optical depth of $\tau_0 = 1 \times 10^7$ and $a = 4.72\times 10^{-4}$ with $x_s = 0.0$. The top panel is linear scale, the middle panel is log scale with $|\rm H_{bc}|$ shown instead of $\rm H_{bc}$, and the bottom panel is the residual of each solution with Monte Carlo.} 
    \label{fig:sol_mc_residual_0}
\end{figure}

We now compare each of the previously-discussed solutions for surface flux to the Monte Carlo results. The spectrum $P(x)$ is defined as the specific luminosity at the surface divided by the source luminosity, or
\be \label{eq:prob_spectrum}
P(x) = \frac{16\pi^2R^2H(R, x)\Delta}{L}.
\ee
This is normalized as $\int P(x)dx = 1$. Since $H(r, x)$ is per $d\nu$, a factor of $\Delta$ gives the expression the correct units.

@@

The size of $H_{\rm bc}$ is dependent on $\tau_0$. $H_{\rm bc}$ is significant even at $\tau_0 {\sim} 10^7$ where the $H_0$ solution is expected to perform well, i.e., photons are pushed further out into the wing where the simplifying assumptions made in the derivation of the differential equation are a better approximation.

In Figure \ref{fig:sol_mc_tau} we show the solutions alongside Monte Carlo for $\tau_0=10^5, 10^6$, and $10^7$. From Eq. \ref{eq:hbc_scaling}, the size of the term $H_{\rm bc}$ should become smaller with larger optical depths, following a $(a\tau_0)^{-1/3}$ scaling. Indeed, agreement between Eq. \ref{eq:H0surf} and the Monte Carlo points in Fig. \ref{fig:sol_mc_tau} improves as $\tau_0$ increases, with $H_{\rm bc}$ providing a fractionally smaller correction to $H_0$. One factor of $(a\tau_0)^{1/3}$ has been scaled out of the x-axis such that the peaks of the distributions are horizontally aligned. This scaling has also been applied to the y-axis to preserve normalization of the escape probability. At lower $\tau_0$, the scattering of photons within the Doppler core of the line becomes important, but our analytic solution does not include this effect. The effects of line core scattering can be seen in the Monte Carlo data for $\tau_0=10^5$ and, to a lesser extent, $
\tau_0=10^6$.

 \begin{figure}
    \centering
    \includegraphics{xinit_threepanel.pdf}
    \caption{The same as Figure \ref{fig:sol_mc_residual_0}, but for $\rm x_s = 0.0, 6.0, 12.0$. The optical depth at each of these source frequencies is $\tau_s = 10^7, 77,$ and $19$, respectively.} 
    \label{fig:sol_mc_xinit}
\end{figure}

 \begin{figure}
    \centering
    \includegraphics{xinit_threepanel_tau1e6.pdf}
    \caption{The same as Fig. \ref{fig:sol_mc_xinit}, but at a lower optical depth $\tau_0 = 10^6$. The shift $x_s$ is a much larger fraction of the distance to the spectral peak $(a\tau_0)^{1/3}$, and thus the asymmetry in the spectrum is much larger. The optical depth at the source frequency is $\tau_s = 10^6$, $7.7$, and $1.9$ for $x_s=0.0, 6.0,$ and $12.0$, respectively. } 
    \label{fig:sol_mc_xinit_lowtau}
\end{figure}

@@

Figure \ref{fig:sol_mc_xinit} shows calculations performed for $x_s = 0.0, 6.0$, and $12.0$ and $\tau_0=10^7$. The asymmetry of the spectrum is slight for $\rm x_s = 6.0$ where $\tau(x_s)=77$, but is larger for $\rm x_s=12.0$ outside the line core where $\tau(x_s)=19$. It is seen here that the difference between the Monte Carlo data and $H_0$ becomes larger as $\rm x_s$ increases. Thus, for large $|x_s|$, inclusion of $H_{\rm bc}$ is more important.

Figure \ref{fig:sol_mc_xinit_lowtau} shows emission away from line center at the same values of $x_s$ as in Figure \ref{fig:sol_mc_xinit}, but for $\tau_0=10^6$ rather than $10^7$. The difference between the left and right side of the escaping spectrum is now substantial since $(a\tau_0)^{-1/3}$ increased by a factor of ${\sim}2$. It is clear from the figure that as $x_s$ extends further into the wing, the spectrum becomes more strongly peaked in frequency. Additionally, since the sphere is increasingly optically thin on the wing, we expect there to be a stronger disagreement with the Monte Carlo as the analytic solution assumed large optical depths.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{xinit.pdf}
    \caption{$H_{\rm bc}$ vs. frequency. Here $P(x) = 16\pi^2R^2H_{\rm bc}(R, x)\Delta/L$, as in Eq. \ref{eq:prob_spectrum}. Thicker lines indicate larger optical depth labeled by $\tau_0=10^5, 10^7,$ and $10^9$, and darker lines indicate $\sigma_s$ further from line center, labeled by $\sigma_s=0.0, \tau_0, 2\tau_0$. The location of the core-wing boundary $x_{cw}$ is shown with green vertical lines of increasing thickness for each $\tau_0$. The axes are scaled to show that the size of the corrective factor agrees with the predicted $(a\tau_0)^{-1/3}$ scaling.}
    \label{fig:xinit}
\end{figure}

Figure \ref{fig:xinit} shows how the boundary term $H_{\rm bc}$ scales with both $\tau_0$ and $x_s$. In this figure, $\sigma_s$ is shifted by integer multiples of $\tau_0$ (Eq. \ref{eq:change_of_variables}) such that the source falls near the peak of the spectrum for each $\tau_0$. For clarity, only the $x > 0$ side of the spectrum is shown. From Eq. \ref{eq:hbc_scaling}, it is expected that the fractional size of $H_{\rm bc}$ relative to $H_0$ should become smaller with larger optical depths following $(a\tau_0)^{-1/3}$. This factor has been scaled out of the figure such that solutions for different $\tau_0$ and the same $\sigma_s$ should show close agreement in scale on the figure's vertical axis if the relation holds. Indeed, the scaled solutions converge as $\tau_0$ becomes larger, indicating agreement with the $(a\tau_0)^{-1/3}$ scaling. The remaining discrepancy present in the vertical axis for fixed $\sigma_s$ results from $x_{\rm peak}$ becoming close to $x_{\rm cw}$; at lower $\tau_0$, this causes the line profile approximation in the wing (Eq. \ref{eq:app:line_profile_wing}) to break down. From this, we conclude that the errors introduced by the incorrect separation of variables in \cite{1973MNRAS.162...43H}, \cite{1990ApJ...350..216N}, \cite{2006ApJ...649...14D} and others are indeed proportional to $(a\tau_0)^{-1/3}$.

@@

In order to understand how long it takes for the photons to escape the uniform sphere of gas we must reintroduce the time dependence of the diffusion equation, which was ignored in the steady-state calculations in Section \ref{sec:steadystate}. To obtain the radiative intensity $I=dE/(dAdtd\Omega d\nu)$ on timescales comparable to the light-crossing time $t_{\rm lc} = R/c$, the time-dependent response to a delta function impulse is found. This allows the distribution of photon escape times (the ``wait time distribution'') to be characterized. For simplicity, a $J=0$ boundary condition will be used in the following derivations, which is a rough approximation for $a\tau_0 \gg 1$.
